Prediction of 2024 US election ...

### 2024-11-02 paper/paper.qmd
colin code update3

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
library(arrow)
library(rstanarm)
library(brms)
```

# Introduction {#sec-intro}
Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States. Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result. As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts. However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data. A more adaptable model is required to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.

We build a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics. Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.

In this paper, we predict the 2024 us election result with the hybrid election forcasting model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that uses Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S. presidential election.

@@

For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
- **response variable:** pct(pct: The percentage of the vote or support that the candidate received in the poll) \
- **numeric predictor:**\
 sample size(sample_size: The total number of respondents participating in the poll) \
 timegap(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
 pollscore(A numeric value representing the score or reliability of the pollster in question) \
- **categorical predictor**
state(The U.S. state where the poll was conducted or focused) \
methodology(The method used to conduct the poll)\

## Explore the data

@@

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned US election data 
#| echo: false
#| warning: false
cleaned_data = read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))

@@

## Basic Model
```{r}
just_harris_data = Harris_data |> na.omit()
model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology, data = just_harris_data)
summary(model_MLR)
just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))

@@

@@

bayesian_model_1 <- stan_glmer(
  formula = formula,
  data = just_harris_data,
  family = gaussian(),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)
pp_check(bayesian_model_1)
summary(bayesian_model_1)

# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)

@@

# Transform Biden and Trump’s vote shares to fit a beta distribution
# Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
total_votes <- 100
biden_shape1 <- 51.3 / total_votes * 10
biden_shape2 <- (1 - 51.3 / total_votes) * 10
trump_shape1 <- 46.8 / total_votes * 10
trump_shape2 <- (1 - 46.8 / total_votes) * 10

# Set up priors using beta distributions based on 2020 vote shares
# prior <- c(
#   set_prior(paste0("beta(", biden_shape1, ", ", biden_shape2, ")"), class = "b", coef = "biden"),
#   set_prior(paste0("beta(", trump_shape1, ", ", trump_shape2, ")"), class = "b", coef = "trump")
# )

prior <- c(
  set_prior(beta(biden_shape1, biden_shape2), class = "b", coef = "biden"),
  set_prior(beta(trump_shape1, trump_shape2), class = "b", coef = "trump"))

# Define the model formula

formula1 <- pct ~ pollscore + days_taken_from_election + methodology

# Run the Bayesian model
model <- brm(
  formula = formula1,
  data = baye_model_data,
  family = categorical(),
  prior = prior,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  cores =4
)

# Summarize the model
summary(model)

@@

# Methodology of YouGov
YouGov's methodology documentations are seperated in two articles. The article by @bailey_how_2024 documents the methodology of the 2024 election projection, while the webpage on @yougov_methodology_nodate documents the general methodology of YouGov's prediction.

## Population, Frame, and Sample
As @bailey_how_2024 stated, the population covered by YouGov's MRP model is everyone in the national voter file, whether or not they belong to YouGov's panel. The national voter files are digital database built by commercial organizations with public government records of voters, as explained by @desilver_q_2018. Voter files indicates whether asomeone voted in a given election, thus YouGov's population covers all voters in previous US elections.

YouGov’s sampling frame consists of its online panel members. These members are part of the SAY24 project, a collaboration between Stanford, Arizona State, and Yale Universities, as stated by @bailey_how_2024. YouGov collect information on respondents when they join their panel before they are invited to participate in the survey.

YouGov select the sample from the sampling frame based on their ability to match characteristics of the population of interest. YouGov interviews nearly 100,000 people in the first set of estimates. For the second set of estimates, YouGov didn't just start over with a new sample. They took the initial data from August and September and updated it with responses from more than 20,000 additional registered voters who were re-interviewed in late September and early October.

## Sample Recruitment
Panelists are recruited through various online channels, including advertisements and partnerships with websites [@yougov_methodology_nodate]. They must provide demographic details upon joining, which helps in selecting representative samples for each survey. When respondents complete a survey, they are awarded points that can be exchanged for money.

## Sampling Approach and Trade-offs
YouGov uses non-probability sampling due to the compensation, an approach where not every individual has an equal chance of selection [@yougov_methodology_nodate]. This method allows quick and cost-effective data collection. However, as @yougov_methodology_nodate writes the panelists must have an internet connection to participate. YouGov state that there is 95% of us population with internet access, thus the sample may be less representative of certain hard-to-reach populations, such as individuals with very slow internet access or without internet access.

## Non-response Handling
YouGov apply statistical weighting to adjust for the differences between the sample and target population. The weight is based on demographic characteristics such as age, gender, race and presidential vote [@yougov_methodology_nodate]. Additionally, quality control measures exclude unreliable responses to improve data accuracy. The respondents are offered a small incentive to decrease the non-response and increase participation.

## Strengths and Weaknesses of the Questionnaire
YouGov's surveys are conducted online, which is very efficient for the respondents, and responses are weighted to enhance representativeness. The pollster can recruit a large amount of panelists because of the online format. Combining with online tracking technologies, the metadata provided by ther panelists can be verified easily.

As a non-probability sample, it might miss certain demographic groups not covered by the online population. While weighting improves accuracy, it cannot fully substitute the randomization found in probability sampling​. Additionaly, the categories in the surbey is oversimplified with bias. For instance, in the poll result published by YouGov, gender is divided into Male and Female. Race is divided into White, Black, Hispanic and Other. This indicates a lack of representation.

>>>>

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
library(brms)
library(rstanarm)
library(gridExtra)
```

```{r}
cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
```

# Introduction {#sec-intro}
Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States. Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result. As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts. However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data. This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.

This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics. Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.

In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S. presidential election.

@@

For the data we used in this analysis about the polling result for Kamala Harris and Donalad Trump in 2024 USA president election. \
- **response variable:** `pct`(pct: The percentage of the vote or support that the candidate received in the poll) \
- **numeric predictor:**\
 `sample size`(sample_size: The total number of respondents participating in the poll) \
 `timegap`(the time gap between the poll start date and the real election date i.e timegap = real US election date - poll start date) \
 `pollscore`(A numeric value representing the score or reliability of the pollster in question) \
- **categorical predictor**
`state`(The U.S. state where the poll was conducted or focused) \
`methodology`(The method used to conduct the poll)\

## Explore the data
```{r}
#| label: tbl-explore-data
#| echo: false
#| warning: false
a <- ggplot(data=cleaned_data, aes(x=sample_size, y=pct)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'sample size', y='pct', 
       title = 'pct vs sample size')

b <- ggplot(data=cleaned_data, aes(x=pollscore, y=pct)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'pollscore', y='pct', 
       title = 'pct vs pollscore')

c <- ggplot(data=cleaned_data, aes(x=methodology, y=pct)) + 
  geom_boxplot() + 
  labs(x = 'wind speed', y='pct', 
       title = 'pct vs methodology')

d <- ggplot(data=cleaned_data, aes(x=days_taken_from_election, y=pct)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'max gust speed', y='pct', 
       title = 'pct vs days taken from election')

e <- ggplot(data=cleaned_data, aes(x=state, y=pct)) + 
  geom_boxplot() + 
  labs(x = 'total snow', y='pct', 
       title = 'pct vs state')

grid.arrange(a, b, c, d, e, nrow=2)
```

- pct vs sample size: This scatter plot shows the pct against the sample size, with a fitted trend line indicating a slight positive relationship. The data points are denser for lower sample sizes, suggesting that smaller sample sizes are more common in the dataset.

- pct vs pollscore: This scatter plot illustrates the pct against pollscore. The fitted trend line suggests a weak negative relationship between pollscore and pct. The points are scattered without a strong linear pattern.

- pct vs methodology: A boxplot comparing pct for different polling methodologies. The pct distribution varies across methodologies, with some showing greater spread or median differences. This suggests that the polling methodology may influence pct outcomes.

- pct vs days taken from election: A scatter plot displaying pct versus the number of days before the election. The trend line indicates a slight negative relationship, suggesting that as the election date approaches, pct may decrease slightly.

- pct vs state: A boxplot depicting pct across different states. The pct distribution varies by state, with some states showing wider variability or different median values, implying state-specific effects on pct.

```{r}
# numeric_data <- cleaned_data[sapply(cleaned_data, is.numeric)]
numeric_data = cleaned_data |> select(pct, sample_size, pollscore, days_taken_from_election)

# Create the pairs plot
pairs(numeric_data)
```

The pairs plot displays scatter plots of four numeric variables (`pct`, `sample_size`, `pollscore`, and `days_taken_from_election`) to visualize their relationships. The data shows clustering, particularly in `pct` versus `sample_size`, suggesting potential heteroscedasticity. The `sample_size` variable is skewed towards lower values, while `pollscore` and `days_taken_from_election` have a more even spread, though `pollscore` shows central clustering. No strong linear relationships are immediately apparent between the variables, indicating that correlations are likely weak.

@@

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned US election data 
#| echo: false
#| warning: false

@@

## Basic Model
```{r}
#| echo: false
#| warning: false
just_harris_data = Harris_data |> na.omit()
lm_model1 = lm(pct ~ pollscore, data = just_harris_data)
predictions = predict(lm_model1, just_harris_data)
summary(lm_model1)

ggplot(just_harris_data, aes(x = pct, y = predictions)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()

```{r}
#| echo: false
#| warning: false
just_harris_data = Harris_data |> na.omit()
model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology + sample_size + state, data = just_harris_data)
summary(model_MLR)
just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))

@@

```{r}
#| echo: false
#| warning: false
ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()
```

@@

bayesian_model_1 <- stan_glmer(
  formula = formula,
  data = just_harris_data,
  family = gaussian(),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)
pp_check(bayesian_model_1)
summary(bayesian_model_1)
```

This plot displays the results of a posterior predictive check (PPC) for a Bayesian model using the pp_check() function.

The observed data (y) follows a distinct, symmetrical distribution centered around a specific value, suggesting a well-defined peak with tapering on both sides. The replicated distributions (y_rep), shown as multiple thin lines, generally align with the shape of the observed distribution, indicating that the Bayesian model has captured the main characteristics of the data. However, the variability in the y_rep lines highlights the degree of uncertainty inherent in the model’s predictive capability.

The fact that the y_rep curves closely match the overall pattern of the actual y suggests that the model performs reasonably well in replicating the observed data. Minor discrepancies or deviations between the y and y_rep might imply areas where the model could be fine-tuned or adjusted to improve accuracy. Overall, this PPC indicates that the model provides a decent fit to the data, with some variability accounted for in the predictive samples.

```{r}
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
```

This plot displays the estimated posterior distribution for the intercept parameter in the Bayesian model. The intercept represents the baseline level of support, expressed as a percentage, for Kamala Harris in the U.S. election data used in the analysis. The plot features a point estimate (depicted by the central dot) that indicates the mean or median of the posterior distribution for the intercept, and a horizontal line showing the 95% credible interval, which signifies the range within which the true value of the intercept is likely to fall with 95% probability.

The range of this credible interval spans from approximately 50 to 54 percent, suggesting that the model estimates the baseline level of support for Kamala Harris to be around this range. The relatively narrow width of the interval implies a certain degree of confidence in the estimate, indicating that the data used in the model provided a clear signal for the intercept's value.

In conclusion, the estimated intercept, which represents the baseline percentage of support for Kamala Harris in the U.S. election data analyzed, is centered around 52%, with a credible interval spanning approximately from 50% to 54%. This suggests that, according to the Bayesian model, the underlying support for Kamala Harris.

# # Transform Biden and Trump’s vote shares to fit a beta distribution
# # Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
# total_votes <- 100
# biden_shape1 <- 51.3 / total_votes * 10
# biden_shape2 <- (1 - 51.3 / total_votes) * 10
# trump_shape1 <- 46.8 / total_votes * 10
# trump_shape2 <- (1 - 46.8 / total_votes) * 10

# # Set up priors using beta distributions based on 2020 vote shares
# prior <- c(
#   set_prior(paste0("beta(", biden_shape1, ", ", biden_shape2, ")"), class = "b", coef = "biden"),
#   set_prior(paste0("beta(", trump_shape1, ", ", trump_shape2, ")"), class = "b", coef = "trump")
# )

# # Define the model formula
# formula <- pct ~ pollscore + days_taken_from_election + methodology

# # Run the Bayesian model
# model <- brm(
#   formula = formula,
#   data = baye_model_data,
#   family = categ0orical(),
#   prior = prior,
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   cores =4
# )

# # Summarize the model
# summary(model)

@@

@@
